---
title: "si618hw6_report_chongli"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Load data

First load the provided TSV data file using the **read.table()** function. Then print the first 15 rows of the data frame.

```{r echo=FALSE, fig.width=7}
country <- read.table("countrydata_withregion.tsv",head=TRUE,sep = "\t",quote = "", fileEncoding="windows-1252")
head(country,n=15)
```

## Step 2: Scatter plot of log transformed data 

Compute the natural logarithm of the area and the population of each country to produce a scatter plot using the plot(), qplot() or ggplot() functions. The following chart is created using **qplot()**.

```{r echo=FALSE, fig.width=7}
logArea <- log(country$area)
logPop <- log(country$population)
library(ggplot2)
qplot(logArea,logPop, xlab="log(area)",ylab="log(population)")
```

## Step 3: Data aggregation by region

Sum the areas and populations of all countries by region using the **aggregate()** function, respectively. Then create two pie charts using the **pie()** function showing the areas of regions and populations of regions, respectively. The following charts are created using **pie()**.

```{r echo=FALSE, fig.width=7}
aggArea <- aggregate(country$area,by=list(region = country$region), sum)
aggPop <- aggregate(as.numeric(country$population),by=list(region = country$region), sum)
pie(aggArea$x,labels=aggArea$region,main="Area of Regions")
```

```{r echo=FALSE, fig.width=7}
pie(aggPop$x,labels=aggPop$region,main="Population of Regions")
```

## Step 4: Visualization of Population per sq km of Region

Create a new data frame to hold the population per sq km of each region using the **data.frame()** function. Then sort the data frame by population per sq km in decreasing order with the help of the **order()** function. Finally, the following chart is created using **ggplot()**.

```{r echo=FALSE, fig.width=7}
popPerKm <- aggPop$x / aggArea$x
newFrame <- data.frame(popPerKm,region=aggPop$region)
newFrame <- newFrame[order(newFrame$popPerKm,decreasing = TRUE),]
ggplot(newFrame, aes(x=region,y = popPerKm)) +geom_bar(stat = "identity")+theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(x="region",y="Population per sq km of Regions")
```

